---
title: "Reproducible Research PA 1"
output: 
  html_document: 
    keep_md: yes
---

## Loading and Preprocessing the Data

Set woring directory, load the data and transform the date variable

```{r,echo=TRUE}
RR <- read.csv("activity.csv",header = TRUE)
RR$date <- as.Date(RR$date)
```

## What is mean total number of steps taken per day?

Calculate total number of steps taken each day and make a histogram
```{r, echo=TRUE}
library(ggplot2)
RR1 <- complete.cases(RR)
RR2 <- RR[RR1,]
stepsperday <- aggregate(RR2$steps,by=list(RR2$date),FUN=sum)
names(stepsperday)[1] <- "Date"
names(stepsperday)[2] <- "Steps"
ggplot(stepsperday,aes(x=Steps)) +
  geom_histogram(binwidth = 2000,fill="red",col="black") +
  labs(title="Histogram of Total Steps Taken Per Day",y="Frequency")

```

Mean and Median of total number of steps taken each day
```{r,echo=TRUE}
mean(stepsperday$Steps)
median(stepsperday$Steps)
```

## What is the average daily activity pattern?

Time series plot of 5 minute intervals and the average numer of steps taken
```{r, echo=TRUE}
avgstep <- aggregate(RR2$steps,by=list(RR2$interval),FUN=mean)
names(avgstep)[1] <- "Interval"
names(avgstep)[2] <- "Steps"

plot(avgstep$Interval,avgstep$Steps,type = "l",xlab="Intervals",ylab="Average Steps Taken",main="Average Steps Taken Per Interval")
```

Maximum number of steps taken on average among the 5 minute intervals
```{r,echo=TRUE}
avgstep[which.max(avgstep$Steps),1]
```

## Imputing missing values
Calculate total number of NA rows
```{r,echo=TRUE}
RR2 <- RR[!RR1,]
nrow(RR2)
```
Strategy for filling in NA values. I used the mean for each interval.
```{r,echo=TRUE}
for (i in 1:2304) {
  RR2[i,1] <- avgstep[avgstep$Interval==RR2[i,3],2]
}

```
Create new dataset with NA filled in
```{r,echo=TRUE}
RR3 <- RR[RR1,]
RR4 <- rbind(RR3,RR2)
```
Histogram of the total number of steps taken each day with NA filled in
```{r,echo=TRUE}
stepseachday <- aggregate(RR4$steps,by=list(RR4$date),FUN=sum)
colnames(stepseachday) <- c("Date","Steps")
ggplot(stepseachday,aes(x=Steps)) +
  geom_histogram(binwidth = 2000,color="black",fill="red") +
  labs(title="New Total Number of Steps Taken Each Day",y="Frequency")
```

Mean and Median on the new dataset. They are the same this time, though mean is the same as the previous one.
```{r,echo=TRUE}
mean(stepseachday$Steps)
median(stepseachday$Steps)
```

## Are there differences in activity patterns between weekdays and weekends
Create a new factor called Week indicating whether it is weekday or weekend and make a panel plot
```{r,echo=TRUE}
RR4$Week <- NA
for (i in 1:17568) {
  RR4[i,4] <- ifelse(weekdays(RR4[i,2]) =="Saturday"|weekdays(RR4[i,2])=="Sunday","weekend","weekday")
}
RR4$Week <- as.factor(RR4$Week)
RR41 <- RR4[RR4$Week=="weekday",]
RR41 <- aggregate(RR41$steps,by=list(RR41$interval),FUN=mean)
colnames(RR41) <- c("Interval","Steps")

RR42 <- RR4[RR4$Week=="weekend",]
RR42 <- aggregate(RR42$steps,by=list(RR42$interval),FUN=mean)
colnames(RR42) <- c("Interval","Steps")

par(mfrow=c(2,1))
plot(RR41$Interval,RR41$Steps,type="l",main = "Weekday Average Steps Taken Per Interval",xlab="Interval",ylab = "Steps")
plot(RR42$Interval,RR42$Steps,type="l",main = "Weekend Average Steps Taken Per Interval",xlab = "Interval",ylab = "Steps")

```
